Introducing Distributed Dynamic Data-intensive (D3) Science: Understanding Applications and Infrastructure

### 2016-02-14 3dpas_intro.tex
ongoing fixes.

From \S\ref{sec:distTypes} and
\figurename~\ref{fig:figures_application}, we find that applications
with distributed data differ based on the amount or volume of data
distributed, the degree of distribution (scattered over one site or
many sites), and the dynamic aspects associated with distribution,
e.g., latency and relative scheduling flexibility (and thus tolerance
or constraints on data transfer/aggregation).  Applications processing
the largest volumes of data (O(PB)) typically run on specialized data
cyberinfrastructure (e.g., the ATLAS LHC Grid). Applications such as
Fusion and NGS that run on smaller data sets mainly utilize HPC
environments. On the other end of the spectrum are highly dynamic
applications that process sensor data and make near-realtime
decisions based on streaming data, e.g., the power grid and sensor
network applications. These applications process much less data;
often, sensor data is also archived and aggregated for historical
analyses, which leads to higher data volumes. %\katznote{not sure about the `other' part of this...}

@@

%\jhanote{the following does not make sense to
%  me..}  Additional temporal aspects pertain to the stage(s) of an
%application in which data can be distributed.
Because application stages often have a distribution that makes sense
to that stage, and because that distribution may not be appropriate for all other
phases, applications must either use one particular distribution
that may not work well in all phases or transform the data on input to
some stages.  In particular, infrastructural applications
require a specific distribution for the data storage, and thus
the stages that read and process the data will need to make this choice.

As seen, how application data is distributed depends upon which stage
of an application is being considered. In spite of such variation,
{\it data topology} can be used to identify the flow of data through
the application, with nodes representing execution units and edges
representing communication features.
% The overall shape of such a data connectivity pattern gives a sense
% of {\it distributedness}; the thickness of edges gives a sense of
% how data intensive the application is; and the change in the shape
% and thickness (and rate) and number of edges gives a sense of the
% dynamism of the application. \katznote{it seems odd that we describe
%   a picture here but don't actually have any such picture in the
%   paper.}
Additionally, how the application reacts to and handles dynamism is an
important factor and may also indicate that some aspects of the data
topology are unknown.

>>>>

From \S\ref{sec:distTypes} and \figurename~\ref{fig:figures_application}, we
find that applications with distributed data vary in the amount or volume of
data distributed, the degree of distribution (scattered over one site or many
sites), and the dynamic aspects associated with distribution, e.g., latency and
relative scheduling flexibility (and thus tolerance or constraints on data
transfer/aggregation).  Applications processing the largest volumes of data
(O(PB)) typically run on specialized data cyberinfrastructure (e.g., the ATLAS
LHC Grid). Applications such as Fusion and NGS that run on smaller data sets
mainly utilize HPC environments. On the other end of the spectrum are highly
dynamic applications that process sensor data and make near-realtime decisions
based on streaming data, e.g., the power grid and sensor network applications.

% These applications process much less data;
% often, sensor data is also archived and aggregated for historical
% analyses, which leads to higher data volumes. %\katznote{not sure about the `other' part of this...}

@@

%\jhanote{the following does not make sense to
%  me..}  Additional temporal aspects pertain to the stage(s) of an
%application in which data can be distributed.

% Because application stages often have a distribution that makes sense to that
% stage, and because that distribution may not be appropriate for other
% phases, applications must either use one particular distribution that may not
% work well in all phases or transform the data on input to some stages.

Applications must either use one particular distribution, or transform data on
input to some stages. In particular, infrastructural applications require a
specific distribution for the data storage, and thus the stages that read and
process the data will need to make this choice.

As seen, how application data is distributed depends upon which stage of an
application is being considered. In spite of such variation, {\it data topology}
can be used to identify the flow of data through the application, with nodes
representing execution units and edges representing communication features.
Additionally, how the application reacts to and handles dynamism is an important
factor and may also indicate that some aspects of the data topology are unknown.

% The overall shape of such a data connectivity pattern gives a sense
% of {\it distributedness}; the thickness of edges gives a sense of
% how data intensive the application is; and the change in the shape
% and thickness (and rate) and number of edges gives a sense of the
% dynamism of the application. \katznote{it seems odd that we describe
%   a picture here but don't actually have any such picture in the
%   paper.}
